---
output:
  word_document: default
  html_document: default
---

```{r}

```

### Lab 2 - Probability
#### Matt Graham

```{r}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
head(kobe)
```

```{r}
kobe$basket[1:9]
```

### Exercise 1

Streaks of 1 means Kobe made one shot after missing the prior. A streak of 0 means he missed both his current shot and the shot prior. Each streak is independent of the next, explaining the exponential decay in probability of many successions.

```{r}
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
```


```{r}
summary(kobe_streak)
```

### Exercise 2
Observing our binomial distribution, we can see that Kobe's streak lengths in 2009 more mostly distributed between 0 and 1 baskets. He had an equal number for both 2 and 3 shot streaks, and very few 4. We can see our center is at 0 and our Mean is .76, giving us a right skew.


```{r}
subset(kobe, kobe_streak = max(kobe$kobe_streak))
```


## Simulations
```{r}
outcomes <- c("heads", "tails")
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
unfair_results <- table(sim_unfair_coin)
unfair_results
```

### Exercise 3
In the simulation above, we flipped 22 heads and 78 tails. This is very close to our assumed weight of 20/80. If we were to run this simulation 10, 100, 1000+ times, we will get closer to 20/80.

```{r}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 1, replace = TRUE)
sim_basket_res <- table(sim_basket)
sim_basket_res
```

In this one shot, we missed our basket.

### Exercise 4

To adjust our argument, we must add the 45% probability of making an independent shot. Below are our histograms of our independent shooter vs. Kobe. While Kobe may have shot a few more shots, it is not significant. I would use this as a supporting argument in that Kobe did not have "hot hands".

```{r}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))
sim_basket_res <- table(sim_basket)
sim_basket_res
```


```{r}
barplot(table(kobe$basket))
barplot(sim_basket_res)
```


```{r}
m <- 10; n <- 7; k <- 8
x <- 0:(k+1)
phyper(x, m, n, k)
```
