---
output:
  word_document: default
  html_document: default
---

### Assignment 2 - Probability
#### Matt Graham

Kobe and Kobe Streak
```{r}
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")
# head(kobe)
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
summary(kobe_streak)
table(kobe_streak)
```

Simulation base
```{r}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))
sim_basket_res <- table(sim_basket)
sim_basket_res
```

Simulation streak
```{r}
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
summary(sim_streak)
table(sim_streak)
```

### Exercise 1

Oh wow, this is cool! We can see right off the bat how much wider the streaks are in our simulated model. Something we can see are fewer single baskets, giving us a stronger right skew. We can see we typically miss with a streak = 0, but our highest streaks are 5 and 6 shots, respectively. That's way different than what we have for Kobe.

### Exercise 2

I would fully expect the results to be different. Each run - like each shot made - are independent of each other. Not one simulation, or game in our scenario, are the same, regardless of the same player - weights in our simulation. They would be relatively similar - mean, median, and skew will not shift too much. I would expect multiple simulations to level off into insignificant margins of error within our determined probabilities.

### Exercise 3
The main difference between Kobe and our simulated independent shooter lies within the greater probability of higher streaks. Kobe never had a streak greater than 4, we had streaks of 5 and 6 in our simulation. Even then, Kobe had a single streak of 4 compared to multiple that we simulated. Overall - including what we summarized in our lab - I don't think there is enough evidence that Kobe and our simulation have similar shooting patterns.



