---
output:
  word_document: default
  html_document: default
---

### Lab 3 - Normal Distribution
#### Matt Graham

Data gathering and prep
```{r}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")
load("bdims.RData")
head(bdims)
```

Creating subsets between sex-assigned males and females
```{r}
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
```

### Exercise 1

```{r}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)

hist(fdims$hgt, probability = TRUE, ylim = c(0, 0.06))
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")

summary(fdims$hgt)
c(fhgtmean, fhgtsd)
```

```{r}
mhgtmean <- mean(mdims$hgt)
mhgtsd   <- sd(mdims$hgt)

hist(mdims$hgt, probability = TRUE)
x <- 150:210
y <- dnorm(x = x, mean = mhgtmean, sd = mhgtsd)
lines(x = x, y = y, col = "red")

summary(mdims$hgt)
c(mhgtmean, mhgtsd)
```

Overall, the shapes between our sex-assigned male and female demographics are almost identical. This is a strong display of how normalized our heights are as humans. I notice a general theme where sex-assigned males tend to be ~10cm taller across the board. We can also see the variances between the two dats ets - males ~7.1cm and females around 6.5cm. This is significant in that there is a degree of error between the two, which is an important note to keep in mind. Sex-assigned men tend to have more variation in height than sex-assigned women.

### Exercise 2
Absolutely - we have very distinct bell curves. Our medians and means are also very close to each other in our data sets. These tools allow us to understand that our data is likely to be normal. Upon visual inspection, we can confirm this is true.

### Exercise 3

Actual
```{r}
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```

Simulation
```{r}
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm)
qqline(sim_norm)
```

We can see how close our model predicted the heights of sex-assigned females. The plot appears to show even less variation than what we observed. We can see a good majority of our points fall within our line. Something important to note are the opposing curves on the low and high end of our spectrum. This makes sense as these heights tend to present as outliers.

### Exercise 4
The main differences I see between the two are the greater probability for variation the model does not capture. This is an important nuance as we begin asking how we want to approach solving this problem. In short, we can say there is enough evidence that female heights are nearly normal.

### Exercise 5

```{r}
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)

hist(fdims$wgt, probability = TRUE, ylim = c(0, 0.06))
x <- 30:110
y <- dnorm(x = x, mean = fwgtmean, sd = fwgtsd)
lines(x = x, y = y, col = "blue")

summary(fdims$wgt)
c(fwgtmean, fwgtsd)
```

Here we can see our right skew, this is evidenced with our disparate mean and median. While the skew does not appear to be significant off the bat, we must consider the sensitivities around using models and factoring weight predictions. There are too many factors to write this as "senseless".


Simulation on weight
```{r}
sim_norm_wgt <- rnorm(n = length(fdims$wgt), mean = fhgtmean, sd = fhgtsd)
qqnorm(sim_norm_wgt)
qqline(sim_norm_wgt)
```

When running our models, we can see the variation is slightly captured. What stands out to me is there doesn't appear to be as s trong of a skew that I expected.


```{r}
qqnorm(fdims$wgt)
qqline(fdims$wgt)
```

Our QQ plot displays how strong of a skew our data has on what normal is. The bow shape is a tell-tale sign for abnormality. This is much more of what I expected our model to generate. An important factor to keep in mind is being mindful of how people feel about and in their bodies. Folks across the spectrum struggle with things we cannot imagine.

### Exercise 6


What is the probability of a randomly selected, sex-assigned female is heavier than 80kg?
```{r}
1 - pnorm(q = 80, mean = fwgtmean, sd = fwgtsd)
sum(fdims$wgt > 80) / length(fdims$hgt)
```

There is ~2.182% (theoretical) and ~4.231%(empirical) chance that we select a sex-assigned female who is heavier than 80kg.

What is the probability of a randomly selected, sex-assigned female is shorter than 150cm (5'0")?
```{r}
pnorm(q = 150, mean = fhgtmean, sd = fhgtsd)
1 - sum(fdims$hgt > 150) / length(fdims$hgt)
```

There is ~1.153% (theoretical) and ~1.153%(empirical) chance that we select a sex-assigned female who is shorter than 150cm (5'0"). Extremely close!