---
output:
  word_document: default
  html_document: default
---

### Assignment 0
#### Matt Graham

First we execute our handout exercises, then we will wrap up with answers from the assignment

#### Handout
```{r}
source("http://www.openintro.org/stat/data/arbuthnot.R")
head(arbuthnot)
```

```{r}
dim(arbuthnot)
```

```{r}
names(arbuthnot)
```

```{r}
arbuthnot$boys
```

```{r}
plot(x=arbuthnot$year, y=arbuthnot$girls)
```

```{r}
plot(x=arbuthnot$year, y=arbuthnot$girls, type = "l")
```


```{r}
?plot
```

```{r}
arbuthnot$boys + arbuthnot$girls
```

```{r}
plot(x=arbuthnot$year, y=arbuthnot$boys + arbuthnot$girls, type = "l")

```

```{r}
arbuthnot$boys / arbuthnot$girls
```


```{r}
arbuthnot$boys / (arbuthnot$boys + arbuthnot$girls)
```

```{r}
arbuthnot$boys > arbuthnot$girls
```



### Assignment

#### Get data
```{r}
source("http://www.openintro.org/stat/data/present.R")
head(present)
```

#### 1. What years are included in this data set? What are the dimensions of the data frame and what are the variable or column names? 

```{r}
c(min(present$year), max(present$year))
```
The dataset contains data from 1940 to 2002.

```{r}
dim(present)
```

The dataset is 63 rows long and 3 columns wide. Our column/variable names are below.
```{r}
names(present)
```

#### 2. How do these counts compare to Arbuthnot’s? Are they on a similar scale?
Relatively! 63 rows is not significantly less than 82 in the grand scheme of data. 

#### 3. Make a plot that displays the boy-to-girl ratio for every year in the data set. What do you see? Does Arbuthnot’s observation about boys being born in greater proportion than girls hold up in the U.S.? Include the plot in your response.

```{r}
plot(x = present$year, y = (present$boys / present$girls), type="l")
```

We can see the representation of boys has moved closer to that of 50/50. This can be attributed to many factors, mostly that of legislation to improve equitable reporting. We can also consider the number of hospitalized births compared to at-home births, something I am sure was more common in early times. 

So no, we cannot conclude that boys are being born in a greater proportion to girls in the U.S. Especially not significantly.

#### 4. In what year did we see the most total number of births in the U.S.? You can refer to the help files or the R reference card http://cran.r-project.org/doc/contrib/Short-refcard.pdf to find helpful commands

```{r}
present[which.max(present$boys + present$girls),]
```

