
```{r}
library(readxl)
library(reshape2)
library(ggplot2)
library(lmtest)
# library(car)
```



### Getting our data
```{r}
meap <- read_excel('meap00_01.xls')
meap
```


Estimate OLS model: math4 =  f { lunch, log(enroll), log(exppp) }
EXPPP = expenditures per pupil: expend/enroll

```{r}
math_mod <- lm(math4 ~ lenroll + lexppp, data=meap)
summary(math_mod)
```


Obtain the usual standard error, and the robust standard errors. How do they compare?
```{r}
coeftest(math_mod)
```

```{r}
coeftest(math_mod, vcov=hccm)
```

Apply White test for HSK. What is the value of the F-test?  What do you conclude?
```{r}
bptest(math_mod, ~ fitted(math_mod) + I(fitted(math_mod)^2) )
```

Our F-test is 84.435 and our p-value is very small, we can conclude there is HSK.


Obtain the fitted values of math4, call it y.  Obtain the residuals, u.

```{r}
u <- resid(y)
plot(u)
```

We can see our residuals are random and centralized around 0. This means we have no trend in our model's data. This means we can summarize the significance of our variables. 

Our variables result significant across the board. This implies that using log(enroll) and log(exppp) are good to predict the % of students eligible for free or reduced lunch. What's important to note are our R^2 values. They do not capture enough variability to result in accurate predictions.


Run log(u^2) = f { y, y^2}
```{r}
# FGLS: estimation of the variance function
logu2 <- log(u^2)
varreg<-lm(logu2~lunch +log(enroll)+log(exppp), data=meap)
summary(varreg)
```

Let h = exp(y); use h to obtain WLS estimates.  

```{r}
w <- 1/exp(fitted(varreg))
fgls <- lm(math4~lunch +log(enroll)+log(exppp), data=meap, weight=w)
summary(fgls)
```


Are the WLS estimates different from OLS estimates?
Yes, significantly! I'm not sure if I did something wrong here. We can see that lunch is significant for both, even if they have opposite symbols. The intercepts are also significantly different. The R^s for WLS is significantly higher than that for OLS.

Out of WLS and OLS, which one is the more precise?
Based on what we saw, I trust WLS more than OLS. I'm excited to see what other folks got out of this!
