---
title: "DA410_Project2_MattGraham"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

This is our first project, analyzing air pollution, mortality rates, and relevant parameters.

```{r Imports, message=FALSE}
library(nnspat)  # used for dist2full()
library("dplyr")  # used to select numeric datatypes
library("ggplot2")
library(reshape)  # used for melting matricies
```


```{r}

```


### 8.7

For the psychological data in Table 5.1, the discriminant function coefficient 
vector was given in Example 5.5. 
(a) Find the standardized coefficients. 
(b) Calculate ί-tests for the individual variables. 
(c) Compare the results of (a) and (b) as to the contribution of the variables 
to separation of the two groups. 
(d) Find the partial F for each variable, as in (8.26), and compare with the 
standardized coefficients.

```{r}

psych <- read.table("C:/mattgraham93.github.io/school/22_3_DA410/data/T5_1_PSYCH.DAT", header=FALSE)
colnames(psych) <- c('method', 'y1', 'y2', 'y3', 'y4')
as.data.frame(psych)
```



### 8.11 (a and b)

 Using the fish data in Table 6.17, do the following: 
(a) Find the eigenvectors of E_ 1 H . 
(b) Carry out tests of significance for the discriminant functions and find 
the relative importance of each as in (8.13), λ*/ ^ ■ λ^. Do these two 
procedures agree as to the number of important discriminant functions?
```{r}

fish <- read.table("C:/mattgraham93.github.io/school/22_3_DA410/data/T6_17_FISH.DAT", header=FALSE)
colnames(fish) <- c('method', 'y1', 'y2', 'y3', 'y4')
as.data.frame(fish)


```



### 8.15

Carry out a stepwise selection of variables on the fish data of Table 6.17. 

```{r}



```
