---
title: "DA410_Assignment4_MattGraham"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

This is our first project, analyzing air pollution, mortality rates, and relevant parameters.

```{r Imports, message=FALSE}
library(nnspat)  # used for dist2full()
library("dplyr")  # used to select numeric datatypes
library("ggplot2")
library(reshape)  # used for melting matricies
library(klaR)
```


### 8.7

For the psychological data in Table 5.1, the discriminant function coefficient 
vector was given in Example 5.5. 

```{r}
psych <- read.table("C:/mattgraham93.github.io/school/22_3_DA410/data/T5_1_PSYCH.DAT", header=FALSE)
colnames(psych) <- c('gender', 'y1', 'y2', 'y3', 'y4')

# create subsets and remove gender column
male <- psych[psych$gender==1,-1]
female <- psych[psych$gender==2,-1]

as.data.frame(psych)
```

#### (a) Find the standardized coefficients. 

Getting covariance matricies for subsets
```{r}

cov.male <- cov(male)
cov.female <- cov(female)

```

Male covariance matrix
```{r}
as.data.frame(cov.male)
```

Female covariance matrix
```{r}
as.data.frame(cov.female)
```

Calculating pooled variance:
```{r}
male.n <- nrow(male)
male.df <- male.n - 1

female.n <- nrow(female)
female.df <- female.n - 1

psych.n <- nrow(psych)

psych.pooled.var <- ( (male.df*cov.male) %*% (female.df*cov.female)) / psych.n
as.data.frame(psych.pooled.var)
```

```{r}
a <- matrix(c(.5104, -.2033, .4660, -.3097))

a.star <- sqrt(diag(psych.pooled.var)) %*% a
a.star
```
The result is not as expected/given in the book. I expected to return a matrix and was given one single value.

(b) Calculate ί-tests for the individual variables. 

```{r}
lapply(psych[-1], function(x) t.test(x~psych$gender))
```

(c) Compare the results of (a) and (b) as to the contribution of the variables 
to separation of the two groups. 

Given part a did not calculate as expected, I can't determine the differences between the two. I can conclude the rank for part b in important are: y3, y1, y2, then y4.

(d) Find the partial F for each variable, as in (8.26), and compare with the 
standardized coefficients.

```{r}

full.psych <- lm(gender ~y1+y2+y3+y4, data=psych)

reduced.psych.y1 <- lm(gender ~y2+y3+y4, data=psych)
reduced.psych.y2 <- lm(gender ~y1+y3+y4, data=psych)
reduced.psych.y3 <- lm(gender ~y1+y2+y4, data=psych)
reduced.psych.y4 <- lm(gender ~y1+y2+y3, data=psych)

```


y1 ANOVA table
```{r}
anova(reduced.psych.y1, full.psych)
```

y2 ANOVA table
```{r}
anova(reduced.psych.y2, full.psych)
```

y3 ANOVA table
```{r}
anova(reduced.psych.y3, full.psych)
```

y4 ANOVA table
```{r}
anova(reduced.psych.y4, full.psych)
```

Observing our linear models, we can see that our ranks across all 4 reduced models = 3. We can also conclude that y2 is not a significant variable when looking to seek differences between genders. While it may have fallen within our initial scope of "reasonable", it's lack of variability does not make it a good fit for our model.

### 8.11 (a and b)

 Using the fish data in Table 6.17, do the following: 

```{r}

fish <- read.table("C:/mattgraham93.github.io/school/22_3_DA410/data/T6_17_FISH.DAT", header=FALSE)
colnames(fish) <- c('method', 'y1', 'y2', 'y3', 'y4')

method1 <- fish[fish$method==1, -1]
method2 <- fish[fish$method==2, -1]
method3 <- fish[fish$method==3, -1]

method1.bar <- colMeans(method1)
method2.bar <- colMeans(method2)
method3.bar <- colMeans(method3)
y.bar.all <- colMeans(fish[-1])

as.data.frame(fish)
```

Calculate E and H
```{r}
method1.bar.diff <- method1.bar - y.bar.all
method2.bar.diff <- method2.bar - y.bar.all
method3.bar.diff <- method3.bar - y.bar.all

H <- 12 * unname(method1.bar.diff %*% t(method1.bar.diff) 
                 + method2.bar.diff %*% t(method2.bar.diff) 
                 + method3.bar.diff %*% t(method3.bar.diff)
                 )

"compute.within.matrix" <- function(data, mean) { 
  ret <- matrix(as.numeric(0), nrow=4, ncol=4)
  
  for (i in 1:12) { 
      diff <- as.numeric(data[i,] - mean)
      ret <- ret + diff %*% t(diff) 
  }
  return(ret) 
}

E <- compute.within.matrix(method1, method1.bar) + compute.within.matrix(method2, method2.bar) +
compute.within.matrix(method3, method3.bar)

E.H <- solve(E) %*% H
```


Matrix E
```{r}
as.data.frame(E)
```

Matrix H
```{r}
as.data.frame(H)
```


Matrix E*H
```{r}
as.data.frame(E.H)
```

(a) Find the eigenvectors of E^-1 * H . 

```{r}

n <- 12  # total records
k <- 3  # methods of cooking
p <- 4  # dependent variables (judges)

eigen(E.H)

```


(b) Carry out tests of significance for the discriminant functions and find 
the relative importance of each as in (8.13). Do these two 
procedures agree as to the number of important discriminant functions?

```{r}

vals <- eigen(E.H)[1]
eigen_mean <- sapply(vals, mean)
sprintf("Eigenvalue mean: %s", eigen_mean)

sapply(vals, FUN = '/', FUN.VALUE = eigen_mean)

```

Looking at our eigenvalues and eigenvectors, we can conclude our judge, V1, does not obtain agreement between our two procedures. 

### 8.15

Carry out a stepwise selection of variables on the fish data of Table 6.17. 

```{r}

fish.model <- greedy.wilks(fish[-1],fish$method, "lda", niveau = .1)
fish.model
```

When looking at our stepwise selection, we are returned with 2 significant variables. Both y2 and y3 are significant and should be included in our linear models. 

