---
title: "DA420_Project 6_MattGraham"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

Steps needed to get working:
1) Uninstall all versions of R newer than 4.0.5
2) Install R v4.0.5
3) Install RTools 400
4) Install packages; Rstem takes forever. Expect a full 20-30 minutes

```{r}
# install.packages("Rstem", repos = "http://www.omegahat.net/R", type = "source")
# install.packages("devtools")
# install.packages("wordcloud")
# install.packages("SnowballC")
# install.packages("plyr")
# install.packages("ggplot2")
# install.packages("RColorBrewer")
# install.packages("tm")
# require(devtools)
# install_url("https://cran.r-project.org/src/contrib/Archive/sentiment/sentiment_0.1.tar.gz")
# install_url("https://cran.r-project.org/src/contrib/Archive/sentiment/sentiment_0.2.tar.gz")
```

#### Import libraries
```{r Imports, message=FALSE}
library(plyr)
library(ggplot2)
library(wordcloud)
library(RColorBrewer)
library(tm)
library(SnowballC)
library(sentiment)
```

### Getting the data
```{r}
###Get the data
data <- readLines("http://www.r-bloggers.com/wp-content/uploads/2016/01/vent.txt") 
# from: http://www.wvgazettemail.com/
df <- data.frame(data)
textdata <- df$data
```


#### Remove non-essential characters
```{r}
textdata = gsub("[[:punct:]]", "", textdata)
textdata = gsub("[[:digit:]]", "", textdata)
textdata = gsub("http\\w+", "", textdata)
textdata = gsub("[ \t]{2,}", "", textdata)
textdata = gsub("^\\s+|\\s+$", "", textdata)

try.error = function(x)
{
   y = NA
   try_error = tryCatch(tolower(x), error=function(e) e)
   if (!inherits(try_error, "error"))
      y = tolower(x)
   return(y)
}

textdata = sapply(textdata, try.error)
textdata = textdata[!is.na(textdata)]
names(textdata) = NULL
```


### Perform sentiment analysis
```{r}
class_emo = classify_emotion(textdata, algorithm="bayes", prior=1.0)
emotion = class_emo[,7]
emotion[is.na(emotion)] = "unknown"
class_pol = classify_polarity(textdata, algorithm="bayes")
polarity = class_pol[,4]

sent_df = data.frame(text=textdata, emotion=emotion,
                     polarity=polarity, stringsAsFactors=FALSE)
sent_df = within(sent_df,
                 emotion <- factor(emotion, levels=names(sort(table(emotion),
decreasing=TRUE))))
sent_df$polarity
```

```{r}
sent_df$emotion
```


### Plotting our results

#### Emotion category plot
```{r}
ggplot(sent_df, aes(x=emotion)) +
  geom_bar(aes(y=..count.., fill=emotion)) +
  scale_fill_brewer(palette="Dark2") +
  labs(x="emotion categories", y="")
```

#### Polarity plot
```{r}
ggplot(sent_df, aes(x=polarity)) +
   geom_bar(aes(y=..count.., fill=polarity)) +
   scale_fill_brewer(palette="RdGy") +
   labs(x="polarity categories", y="")
```

#### Word cloud
```{r warning=FALSE}
emos = levels(factor(sent_df$emotion))
nemo = length(emos)
emo.docs = rep("", nemo)
for (i in 1:nemo)
{
   tmp = textdata[emotion == emos[i]]
   emo.docs[i] = paste(tmp, collapse=" ")
}
emo.docs = removeWords(emo.docs, stopwords("english"))
corpus = Corpus(VectorSource(emo.docs))
tdm = TermDocumentMatrix(corpus)
tdm = as.matrix(tdm)
colnames(tdm) = emos
comparison.cloud(tdm, colors = brewer.pal(nemo, "Dark2"),
                 scale = c(3,.5), random.order = FALSE,
                 title.size = 1.5)
```


## Analysis
It took a bit to get this working, but I'm so glad I got it right! I left the steps above in case other people experience this problem in the future. 

Most comments are neutral, and a majority are neutral or positive. Relating this to our word cloud, we can see the proportion of positive/negative reflected with negative relating to anger and sadness. surprise and joy are likely attributed with happy, and fear/unknown being neutral.

We can see many words are filed under, "unknown". This can be something addressed and excluded in future visualizations. Or, ideally, revisited and look at things as phrases. Something like, "around" isn't a word of surprise to me. Heart under anger makes no sense either. 

It would be helpful improving the bucketing algorithm. I would like to know why holiday is under surprise and not joy. Why is "aep" part of surprise? Why is it so prominent? "Don't" under joy is also a bit strange.
