```{r}
library(readxl)
library(strucchange) # for chow test
```


The data in 401KSUBS.RAW contains information on netfinancial wealth in dollars (nettfa), age of the survey (age), annual family income in dollars (inc), family size (fsize), and participation in pensions plans.  Use only fsize=1, throughout the following questions (just for this case).

```{r}
beauty <- read_excel('beauty.xls')
beauty
```

Below average folks
```{r}
bel_avg <- subset(beauty, belavg == 1)
bel_avg
nrow(bel_avg)
```

Above average folks
```{r}
abv_avg <- subset(beauty, abvavg == 1)
abv_avg
nrow(abv_avg)
```

Separating by birth-assigned sex

```{r}

male <- subset(beauty, female == 0)
female <- subset(beauty, female == 1)

```


```{r}
male_wage_model <- lm(lwage ~ belavg+abvavg, data=male)
summary(male_wage_model)
```


```{r}
female_wage_model <- lm(lwage ~ belavg+abvavg, data=female)
summary(female_wage_model)
```


3. p-values for below average, across sex-assigned males and females prove to be statistically significant, more so for those for males. An important fact is looking at the base wage, where women earn less than men. So females who are also below average in "beauty standards" earn even less than their below average male counterparts.


4. No - we cannot conclude that sex-assigned females who are above average earn more than women with below average looks. This is determined by looking at our p-value of .54, meaning we have too much spread in our data to conclude this as an effective variable. 


```{r}
female_wage_model_exp <- lm(lwage ~ belavg+abvavg+educ+exper+expersq+union+goodhlth+black+married+south+bigcity+smllcity+service, data=female)
summary(female_wage_model_exp)
```

```{r}
male_wage_model_exp <- lm(lwage ~ belavg+abvavg+educ+exper+expersq+union+goodhlth+black+married+south+bigcity+smllcity+service, data=male)
summary(male_wage_model_exp)
```

```{r}
full_wage_model_exp <- lm(lwage ~ female + belavg + abvavg + educ + exper + expersq + union + goodhlth + black+married + south + bigcity + smllcity + service, data=beauty)
summary(full_wage_model_exp)
```


Chow test
```{r}
male_wage_anova <- anova(male_wage_model_exp)
male_wage_anova
```


```{r}
female_wage_anova <- anova(female_wage_model_exp)
female_wage_anova
```


```{r}
full_wage_anova <- anova(full_wage_model_exp)
full_wage_anova
```


```{r}
k <- 13
N <- 1247

ssrp <- 261.177
ssr1 <- 83.693
ssr2 <- 166.084

f_chow <- ((ssrp - (ssr1 + ssr2)) / (ssr1 + ssr2)) / ((N - (2 * (k + 1))) / k)
f_chow
```

```{r}
qf(.95, 1245, 422, lower.tail = FALSE)
```

Generating our critical F value of 0.8792, and our Chow(F) being 0.0004, we cannot state there is enough evidence for there to be a difference in wage between men and women. This doesn't make sense to me, and am hopeful I did something wrong!

